ODGI: understanding pangenome graphs

### 2021-10-19 manuscript/main.tex
nada
build

We store the graph in a vector of node structures, each of which presents a node-local view of the graph sequence, topology, and path layout.
Expressed in terms of the variation graph $V$, ODGI's core $Node$ structure includes a decoder that maps the neighbors of each node to a dense range of integers.
For a given node $i$ and neighbor $j$, the decoder itself does not store the $id$s of $j$, but rather a compact representation of the relative difference between the node ids: $\delta = Node_i.id - Node_j.id$.
This keeps even the size of the encoding small, per common variation graph property (2).
We define the edges and path steps traversing the node in terms of this alphabet of $\delta$'s.
The structures in algorithm \ref{alg:structs} describes our encoding.

@@

% build
Solves problem of working with big and complex graphs in limited memory.

Figure: encoding model as a cartoon --  HTT exon 1 as tiny example graph (maybe)

>>>>

We store the graph in a vector of node structures, each of which presents a node-local view of the graph sequence, topology, and path layout.
Expressed in terms of the variation graph $V$, ODGI's core $Node$ structure includes a decoder that maps the neighbors of each node to a dense range of integers.
For a given node $i$ and neighbor $j$, the decoder itself does not store the $id$s of $j$, but rather a compact representation of the relative difference between the node ids: $\delta = Node_i.id - Node_j.id$.
This keeps even the size of the encoding small, per common variation graph property (2).
We define the edges and path steps traversing the node in terms of this alphabet of $\delta$'s.
The structures in Algorithm \ref{alg:structs} describes our encoding.

@@

% build
We begin by transforming the ``columnar'' storage model of GFA, in which nodes, edges, and paths are described independently, into the ODGI node-centric encoding with \textit{odgi build}.
This construction step can present a significant bottleneck, in particular as the size of the path set of the graph increases.
The ODGI data structure (Algorithm \ref{alg:structs}) allows algorithms to dynamically construct the graph and its path set to operate in parallel.
In \textit{odgi build}, we initially construct the node vector in a serial operation that scans across the input GFA.
Edges, and then paths, can be added in parallel by obtaining mutex $Node.lock$ for pairs of nodes and adding edges or steps to their corresponding vectors in both nodes.
This parallelism speeds ODGI model construction by many-fold when testing against graphs built by the HPRC (TODO).

%Figure: encoding model as a cartoon --  HTT exon 1 as tiny example graph (maybe)
